## Course: Visual Analytics for Policy and Management
###Bivariate Exercise 2
####Isabella Sun



####Figures:

1. [Exercise 1: Lolipop Plot](#ex1)

2. [Exercise 2: Heat plot in ascending order](#ex2)

3. [Exercise 3: Categorical-Numerical](#ex3)

4. [Exercise 4: Numerical Numerical](#ex4)

##
##

We can start by bringing in the data we will use and libraries
```{r}
link="https://github.com/EvansDataScience/data/raw/master/crime.RData"
load(file = url(link))

library(magrittr)
library(ggplot2)
```
####Exercise 1 Lolipop Plot
create dataframe of what we will plot
```{r}

CrimeCol=table(crime$crimecat,crime$Occurred.DayTime)%>%
         prop.table(margin = 2)%>%   # 2 is % by column
         "*"(100)%>%
         round(3)

df.C=as.data.frame(CrimeCol)
colnames(df.C)=c('Crime','Daytime','Percent')
```
misc elements:

```{r}
title_loli= "What types of crimes are most common during different times of the day?"
caption = "Source: Seattle Open Data Portal - Seattle Police Department "
```
<a id='ex1'></a>
Create Plot
```{r}
base  = ggplot(df.C, aes(x = reorder(Crime, Percent), y = Percent ) ) 

lolliplot1= base + geom_segment(aes(y = 0, 
                                   x = reorder(Crime, Percent), 
                                   yend = Percent, 
                                   xend = Crime), color = "grey50") 
lolliplot2 = lolliplot1 + geom_point()
 
lolliplot3 = lolliplot2 + facet_wrap( ~ Daytime ,nrow = 1) 
lolliplot4 = lolliplot3 + coord_flip() + theme(axis.text.y = element_text(size=4,angle = 45))

lolliplot5 = lolliplot4 + labs(title=title_loli,
                     x ="Crime", 
                     y ="Share of all crimes",
                     caption = caption)

lolliplot6 = lolliplot5 + theme(plot.caption = element_text(hjust = 0),
                      plot.title = element_text(size = 11, face = "bold"),
                      axis.text = element_text(size = 8))


lolliplot6

```

####Exercise 1 Heat Plot
Heat plot in ascending order where intensity goes from yellow to purple
```{r}
title_heat = "What types of crimes are most common during \ndifferent times of the day?"
```
<a id='ex2'></a>
Create Plot
```{r}
base  = ggplot(df.C, aes(x = Daytime, y = reorder(Crime, -Percent), fill = Percent)) 
heat1 = base +  geom_tile() 
heat2 = heat1 +scale_fill_gradient(low = "yellow", 
                                   high = "purple")

heat3 = heat2 + theme_classic() 

heat4 = heat3 + theme(axis.text.x = element_text(angle = 90, vjust = 0.6), 
                      legend.title = element_blank(), #no title for legend
                      legend.position="top", 
                      legend.direction="horizontal",
                      legend.key.width=unit(1, "cm"),
                      legend.key.height=unit(1, "cm")) 

heat5 = heat4 + labs(title=title_heat,
                     caption = caption)

heat6 = heat5 + theme(plot.caption = element_text(hjust = 0),
                      plot.title = element_text(size = 11, face = "bold"),
                      axis.text = element_text(size = 8))

heat6 + labs(y="Crime")
```


####Exercise 3: Categorical-Numerical

Create data frame of what we will plot
```{r}
#before 2000
crimeY2000=crime[(crime$DaysToReport>=365) & (crime$year>=2000),]

crimeY2000$YearsToReport=crimeY2000$DaysToReport/365

crimeYearGG=crimeY2000[complete.cases(crimeY2000$YearsToReport),]

#after 2000
crimeWeek=crime[(crime$DaysToReport<365) & (crime$year>=2000),]

crimeWeek$WeeksToReport=crimeWeek$DaysToReport/7

crimeYearGG2=crimeWeek[complete.cases(crimeWeek$WeeksToReport) &complete.cases(crimeWeek$crimecat),]

```

```{r}
title_2 = "How long does it take for someone to report
different types of crimes? (before 2000)"
```
<a id='ex3'></a>
```{r}
base= ggplot(crimeYearGG,
             aes(x = reorder(crimecat, YearsToReport, FUN = max), # reorder!
                 y=YearsToReport)) 
mins = base + stat_summary(fun.y=min,size=1,
                           geom="line",show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = F,
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median, size=1,
                                    geom="line", linetype='dashed',
                                    aes(group=1,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))

all4 = all3 + labs(title = title_2,
                    x ="Type of Crime", 
                    y = "Weeks to report",
                    caption = caption)

all4 + coord_flip()
```

misc figure elements
```{r}
title_1 = "How long does it take for someone to report
different types of crimes? (after 2000)"
```

Create figure
```{r}
base = ggplot(crimeYearGG2,
              aes(x = reorder(crimecat, WeeksToReport, FUN = max),
                  y=WeeksToReport)) 
mins = base + stat_summary(fun.y=min,size=1.5,
                           geom="line", show.legend = T,
                           aes(group=1,col='Min'))
minsMaxs= mins + stat_summary(fun.y=max,
                              geom="line",size=1,show.legend = T, 
                              aes(group=1,col='Max'))
minsMaxsMd= minsMaxs + stat_summary(fun.y=median,size=1,
                                    geom="line", linetype='dashed',
                                    aes(group=2,col='Median'))
all3=minsMaxsMd + scale_colour_manual(name="Trends",
                                      values=c("blue", "black","red"))

all4 = all3 + labs(title = title_1,
                    x ="Type of Crime", 
                    y = "Weeks to report",
                    caption = caption)

all4+coord_flip()
```

####Exercise 4